---
title: "Bellabeat Case  study"
author: "Sakshi"
date: "2024-09-26"
output: html_document
---

## Introduction

Bellabeat, a high-tech manufacturer of health-focused products for women. Bellabeat is a successful small company. Urška Sršen, cofounder and Chief Creative Officer of Bellabeat, believes that analyzing smart device fitness data could help unlock new growth opportunities for the company.

## Buisness task
* Insight of how consumers are using their smart devices.
* The insights to help guide marketing strategy for the company.

#### Data source: 
<https://www.kaggle.com/arashnic/fitbit>
The dataset was generated by respondents to a distributed survey via Amazon Mechanical Turk between 03.12.2016 and 05.12.2016. And include 18 CSV files.

## Working on the dataset
### Prepare Phase
Installing packages like 'tidyverse', 'ggplot2', 'lubridate', 'dplyr', 'tidyr', 'here', 'skimr', 'janitor' that will help in cleaning, analyzing and plotting our data.


```{r library, message=FALSE, warning=FALSE}
# Loading packages :

library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
library(tidyr)
library(here)
library(skimr)
library(janitor)
```
### Importing datasets
#### Importing dailyActivity_merged.csv
```{r activity import, echo=TRUE, message=FALSE, warning=FALSE}
#daily_activity
daily_activity<- read.csv("C:/Users/saksh/Desktop/case study 2/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/dailyActivity_merged.csv")
head(daily_activity)
```

#### Impoting dailyCalories_merged.csv
```{r calories import, message=FALSE, warning=FALSE}
#calories
calories<-read.csv("C:/Users/saksh/Desktop/case study 2/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/dailyCalories_merged.csv")
head(calories)
```

#### Importing dailyIntensities_merged.csv
```{r intensities import}
#intensities
intensities<- read.csv("C:/Users/saksh/Desktop/case study 2/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/dailyIntensities_merged.csv")
head(intensities)
```

#### Importing heartrate_seconds_merged.csv
```{r heart_rate import}
#heartrate
heartrate<- read.csv("C:/Users/saksh/Desktop/case study 2/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/heartrate_seconds_merged.csv")
head(heartrate)
```

#### Importing sleepDay_merged.csv
```{r sleep import}
#sleep
sleep <- read.csv("C:/Users/saksh/Desktop/case study 2/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/sleepDay_merged.csv")
head(sleep)
```

#### Importing weightLogInfo_merged.csv
```{r weight import}
#weight
weight<- read.csv("C:/Users/saksh/Desktop/case study 2/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/weightLogInfo_merged.csv")
head(weight)
```
So now, we can see that everything were imported correctly.

### Process Phase
And here some cleaning steps I followed:

1.daily_activity, calories and intensities data sets: I did not found any Spelling errors, Misfield values, Missing values, Extra and blank space and duplicated value in the data.'

2.In sleep data set there were 3 duplicates found and removed.

3.In weight data set there were too many missing values found in one column, and I decided to remove that column

there were some data type changes I did 

```{r clean}
#daily_activity
daily_activity$ActivityDate=as.Date(daily_activity$ActivityDate, format="%m/%d/%Y", tz=Sys.timezone())
glimpse(daily_activity)

#intensities
intensities$ActivityDay=as.Date(intensities$ActivityDay, format="%m/%d/%Y", tz=Sys.timezone())

#sleep
sleep$SleepDay=as.POSIXct(sleep$SleepDay, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())

```
Now data sets are cleaned and are ready to start our analysis on them.

### Analyze Phase

#### Looking that data sets are compatible or not

```{r}
#Finding the total number of participants in each data set 
daily_activity%>%
  summarise(total_participants = n_distinct(daily_activity$Id))

n_distinct(calories$Id)
n_distinct(intensities$Id)
n_distinct(heartrate$Id) 
n_distinct(sleep$Id)
n_distinct(weight$Id)
```

As, I can see the data of number of participants in Heart rate data set and Weight data set is very less so, I'll avoid using these data sets to avoid any bias in analysis. 
Now, analyzing data sets to find any type of correlation in the data.

#### Using daily_activity data set for analysis:
```{r daily_activity analysis}
daily_activity %>%  
  select(TotalSteps,TotalDistance,SedentaryMinutes, Calories) %>%
  summary()
```


#### Using intensities data set for analysis:
```{r intensities analysis}
intensities %>%
  select(VeryActiveMinutes, FairlyActiveMinutes, LightlyActiveMinutes, SedentaryMinutes) %>%
  summary()
```

The average sedentary time of participants seems to very high of approx 16.8 hrs and comparatively their active time is very less.

#### Using sleep data set for analysis:
```{r sleep analysis}
sleep%>%
  select(TotalSleepRecords,TotalMinutesAsleep,TotalTimeInBed)%>%
  summary()
```
Participants sleep 1 time for an average of 7.6 hrs and also when the average of participants spends 7.6 hrs of their time on bed then the average of 6.9 hrs they spend on sleeping.


merging data:
```{r merge}
#for most active time
active <- merge(weight, intensities, by="Id", all=TRUE)
active$time <- format(active$Date, format = "%H:%M:%S")
```

### Share Phase

Now let's visualize some key explorations.

#### Relationship between Steps and Sedentary time

```{r}
ggplot(daily_activity)+
  geom_point(mapping = aes(x=TotalSteps, y= SedentaryMinutes))+
  geom_smooth(mapping = aes(x=TotalSteps, y= SedentaryMinutes))+
  labs(title = "Total steps vs sedentary minutes")
```

Through this graph I can see a week negative correlation between total steps taken and sedentary time in a day. This show the more steps you take in a day than there is chances of less sedentary time you spend in a day.

#### Relatonaship between Minutes aspleep and time in bed

```{r}
ggplot(sleep)+
  geom_point(mapping= aes(x=TotalMinutesAsleep, y=TotalTimeInBed))+
  geom_smooth(mapping = aes(x=TotalMinutesAsleep , y= TotalTimeInBed))+
  labs(title = "Total time in bed vs asleep")
```

Through this graph I can see a strong positive correlation between total time in bed and total time asleep in a day. This show that if a person stays more on bed than they sleep more. 

#### Relationship between Calories and steps

```{r}
ggplot(daily_activity)+
  geom_point(mapping = aes(y=Calories, x=TotalSteps))+
  geom_smooth(mapping = aes(y=Calories, x= TotalSteps))+
  labs(title = "Total steps vs Calories")
```

Through this graph I can see a weak positive correlation between total time in bed and total time asleep in a day. This show that if a person take more steps in a day than they are likely to burn more calories in a day. 

#### Looking at intensities over time

```{r}
ggplot(active)+
  geom_line(mapping = aes(x=time, y=VeryActiveMinutes,color="red"))+
  theme(axis.text.x = element_text(angle=90) )+
  labs(title="Total very Active Intensity vs. Time ")
```

As I can see that people are more likely to be active in morning till afternoon and less active during the evening and night in a day.

### Recommandations

After looking at the data and the insights we created 

-As people spends sedentary time a lot so maybe the company should create a marketing strategy focusing on the goal of making people active maybe by including a notification when they spends a lot of sedentary time to motivate them to be more active and spend more time actively. As, the data shows that the company need to market more to the customer segment with a high Sedentary time

-And also we saw that people are more likely to be active since morning till afternoon only so making something to encourage people to maybe go for an evening walk or spending there time more actively

-For customers who want to lose weight, it can be a good idea to control daily calorie consumption and take more steps. And Bellabeat can suggest some ideas for low-calorie healthy food. And also to empower the customers with knowledge about their own health and daily habits.

- On an average the person takes almost 7600 steps per day but according to the CDC, they recommends that most adults aim for 10,000 steps per day that helps in staying active and fit.

Thank you very much for your interest in my Bellabeat Case Study!

And I would appreciate any comments and recommendations for improvement!
